---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

This page will contain all the assignments you submit for the class.




# Assignment 1

**Collaborators: Eliza Epstein. **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.


```{r}
library(datasets)
```

**Answer**: I loaded the library! 

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

```{r}
dat<-USArrests
```

**Answer**: It is nice to rename data because we want to be able to replicate our analysis without messing up the original data.

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=FALSE}
dat$state <- tolower(rownames(USArrests))
```


List the variables contained in the dataset `USArrests`.

```{r}
names(dat)
```
**Answer**: The four variables contained in this dataset are Murder, Assault, UrbanPop and Rape.

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

**Answer**: Murder is a categorical variable, according to the DVB chapter. --- FALSE: It's a quantitative variable.

What R Type of variable is it?

**Answer**: It is a character R Type variable. --- FALSE: It's a numeriical varable. 


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

**Answer**: In this dataset we have the number of arrests per 100,000 residents for each of the following categories: assault, murder and rape in each of the 50 US states in 1973.

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

**Answer**: I chose to make a barplot instead of a histogram because it made more sense to me, considering that the data is categorical. I thought it would be clearer if each bar represented a state, which is why I decided to label the x-axis with the names of the states and the y-axis the number of arrests for murder per 100,000. In the code las=2 simply means that I wanted that state labels to be perpendicular to the x-axis.
```{r}
state.names = row.names(USArrests)
barplot(USArrests$Murder, names.arg = state.names, las = 2, ylab = "Arrest Rate for murder per 100,000", main = "Murder Rate in the United States in 1973")
```


### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
summary(dat$Murder)
```

**Answer**: Min: 0.800, 1st Q: 4.075, Median 7.250, Mean 7.788, 3rd Q 11.250, Max 17.400 ----- The mean is 7.788 and the median is 7.250. On average there were 7.788 (mean) arrests for murder per 100,000 residents in each state in 1973. The median tells us where the middle of the data set is. Since the mean is higher than the median we could say the distribution or arrests for murder per 100,000 in 1973 is positively skewed. A quartile represents 1/4 of the data, for examples the 1st Q is 4.075 and this tells us that 1/4 of the data lies below 4.075 arrests per 100,000. I think R only gives us the 1st and 3rd Quartiles because the median is equal to the 2nd Quartile. 

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r}
state.names = row.names(USArrests)
barplot(USArrests$Assault, names.arg = state.names, las = 2, ylab = "Arrest Rate for assault per 100,000", main = "Assault Rate in the United States in 1973")
```

```{r}
state.names = row.names(USArrests)
barplot(USArrests$Rape, names.arg = state.names, las = 2, ylab = "Arrest Rate for rape per 100,000", main = "Rape Rate in the United States in 1973")
```

```{r}
par(mfrow=c(3,1))
state.names = row.names(USArrests)
barplot(USArrests$Murder, names.arg = state.names, las = 2, ylab = "Arrest Rate for murder per 100,000", main = "Murder Rate in the United States in 1973")

state.names = row.names(USArrests)
barplot(USArrests$Assault, names.arg = state.names, las = 2, ylab = "Arrest Rate for assault per 100,000", main = "Assault Rate in the United States in 1973")

state.names = row.names(USArrests)
barplot(USArrests$Rape, names.arg = state.names, las = 2, ylab = "Arrest Rate for rape per 100,000", main = "Rape Rate in the United States in 1973")

```

```{r, echo = TRUE, fig.width = 5, fig.height = 8}

```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

**Answer**: Par can be used to set different parameters and can allow us to create a single graph from many. 

What can you learn from plotting the histograms together?

**Answer**: By plotting histograms together (or in this case bar charts) we can see patterns more clearly. It becomes easier to compare the rates of arrest per 100,000 residents for each category across the states.
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval = FALSE, fig.width = 7.5, fig.height = 4}
library('maps') 
library('ggplot2') 

ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

**Answer**:The code allows us to make a heat map of arrests for murder per 100,000 in the USA. The first line tells R that we want to make a heat map according to states and that we want to use the data of arrests for murder to fill in each state. The second line creates the map and the last line expands the range to fit in all possible values. 

$$\\[2in]$$




# Assignment 2 

#Carolina Herrera Figueroa 
# CRIM250- Statistics for the Social Sciences
# Assignment 2
# Collaborators: Eliza Epstein

```{r}
setwd("/Users/carolinaherrera/Documents/Assignment 2 - CRIM 250")
```


```{r}
dat <- read.csv(file = 'dat.nsduh.small.1.csv')
```

dat$sexatract
dat$sexatract.nas <-ifelse(dat$sexatract==99, NA, dat$seexatract)

dat$sexatract.nas <- ifelse(dat$sexatract == 99, NA, dat$sexatract)

table(dat$sexatract.nas)

dat$sexatract.nas

sesxnona<-na.omit(dat$sexatract.nas)

table(dat$sexnona)

```{r}
summary(dat)
```

```{r}
head(dat)
```

```{r}
names(dat)
```


```{r}
dim(dat)
```
# The data consists of 171 rows and 7 columns. 

#Problem 2: Variables 
# ANSWER: The variables in this data set are: mjage, cigage, iralcage, age2, sexatract, speakengl, irsex. (mjage) looks at how old the individual was the first time they used marijuana or hashish, (cigage) is have they ever smoked part of a cigarette, (iralcage) looks at how old the individual was when they first tried alcohol, (age2) represents their coded age, (irsex) represents their gender, (sexatract) looks at what best describes their sexual atractment and (speakengl) represents how well the individual speaks English. These are all categorical variables. 

```{r}
names(dat)
```

# What is this dataset about? Who collected the data, what kind of sample is it, and what was the purpose of generating the data?
# This dataset is about drug and alcohol use. It was collected using the National Survey of Drug Use and Health, and this is a random sample of the first 1000 cases of the survey. The purpose of generating the data was to perhaps be able to study the relationship between drug use and other variables such as age or gender. I think the results could perhaps be used to allocate resources to individuals or groups at highest risk of an early onset of drug use.  

-----------------

#Problem 3: Age and gender 
# What is the age distribution of the sample like? Make sure you read the codebook to know what the variable values mean.
# *Answer*: The age distribution of the sample is skewed with most individuals being between the ages of 35-49 years old, where the peak and the mean of the data is located, as shown by Figure 1. The lowest age of a respondent was 15 years old. I think those over 65 years old are also underrepresented in the sample. 

```{r}
summary(dat)
```


```{r}
counts <- table(dat$age2)
barplot(counts, main = "Fig 1. Age of Participants", xlab="Code for Age")
```


# Do you think this age distribution representative of the US population? Why or why  not?
# ANSWER: I think this age distribution is not representative of the US population because it suggests that the population is mostly comprised of individuals between the age of 35-49 years old. 

# Is the sample balanced in terms of gender? If not, are there more females or males?
# ANSWER: The same seems to be relatively balanced in terms of gender, but there are slightly more males than females. The bar chart (Figure 2) shows us that the bar for code "1" which represents males in the codebook is taller, meaning our sample has more males. 

```{r}
counts <- table(dat$irsex)
barplot(counts, main = "Fig 2. Gender of Participants", xlab="Code for Gender")
```


# Use this code to draw a stacked bar plot to view the relationship between sex and age. What can you conclude from this plot?
# ANSWER: This plot shows us that for most age categories an equal number of males and females were surveyed, with the exception of a few. It seems that the younger the participants were the more likely it was that the sample consisted of mostly males or mostly females, but it was not even. For example, for age categories 6 and 7, which translates to the respondents being 17 or 18 years old, the data is made up of males only. Starting at age category 11 we see a sample more evenly split among males and females. 

```{r}
tab.agesex <- table(dat$irsex, dat$age2)
barplot(tab.agesex,
        main = "Fig 3. Relationship between Sex and Age",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.agesex),
        beside = FALSE) # Stacked bars (default)
```



## Problem 4: Substance use

# For which of the three substances included in the dataset (marijuana, alcohol, and cigarettes) do individuals tend to use the substance earlier?
# Answer: Individuals tends to use alcohol earlier, it was reported being consumed as early as 16 years old, compared with 18 years old for marijuana/hashnish and 21 years old for cigarettes. We can also see that 25% of the respondents were under the age of 29 when they tried alcohol for the first time, compared to 25% being below 34 years old the first time they tried marijuana/hashish and 25% of those who respondents who were under 49 years old when theey started smoking cigarettes. Alcohol is also the only substance in our sample that respondents tried for the first time under the age of 18.

```{r}
counts <- table(dat$mjage)
barplot(counts, main = "Fig 4. Age when First Used Marijuana/Hashish", xlab="Age")
```


```{r}
counts <- table(dat$cigage)
barplot(counts, main = "Fig 5. Age when first started Smoking Cigarettes Every day", xlab="Age")
```


```{r}
counts <- table(dat$iralcage)
barplot(counts, main = "Fig 6. Age When First tried Alcohol", xlab="Age")
```


--------------------------
## Problem 5: Sexual attraction

# What does the distribution of sexual attraction look like? Is this what you expected?
# ANSWER: The distribution of sexual attraction is skewed, with a great majority of individuals reporting as being attracted only to the opposite sex. I was not surprised to see such a high number of individuals reporting to be heterosexual, but rather the difference between those who reported to be heterosexual and those who reported to be homosexual. I think we should keep in mind that it's possible that some individuals may have chosen to not report accurately out of fear of embarassment.  

```{r}
counts <- table(dat$sexatract)
barplot(counts, main = " Fig 7:Sexual Attraction", xlab="Code for Sexual Attraction")
```


# What is the distribution of sexual attraction by gender? 
# Answer: The distribution of sexual attraction by gender is interesting because those who identify as strongly being heterosexual (1 on the codebook) are mostly males, while those who identify as being "mostly heterosexual" are mostly female. I thought this was interesting to see because it may be showing us the importance of the wording of the question since males were more likley to choose the option that strongly asserted their sexuality. In our sample of 1,000 individuals, those who identified as bisexual were only female. There was an even split between genders among those who identified as being attracted only to the same sex.

```{r}
tab.sexatract <- table(dat$irsex, dat$sexatract)
barplot(tab.sexatract,
        main = "Fig 8: Relationship between Gender and Sexual Attraction",
        xlab = "Code for Sexual Attraction", ylab = "Frequency",
        legend.text = rownames(tab.sexatract),
        beside = FALSE) # Stacked bars (default)
```





## Problem 6: English speaking

# What does the distribution of English speaking look like in the sample? Is this what you might expect for a random sample of the US population?
# ANSWER: The distribution of English speaking looks very skewed in the sample, with the majority of individuals speaking English very well. The barplot shows us that over 150 individuals speak English very well while about 10 speak English not well. This is not what I would expect of a random sample of the US population, I expected to see a greater number of individuals who spoke English "well" and "not well" considering that English is not the nativie language for a great portion of the population. I didn't expect the data to be as skewed at it is.    

# Are there more English speaker females or males?
# ANSWER: There are more English speaking males, and in our sample none fall under the category of speaking English "not well", represented by the bar labeled "3" on our plot. 

```{r}
counts <- table(dat$speakengl)
barplot(counts, main = "Figure 9: How Well do Participants Speak English", xlab = "Categories")
```


```{r}
tab.sexengl <- table(dat$irsex, dat$speakengl)
barplot(tab.sexengl,
        main = "Figure 10: English Proficiency by Gender",
        xlab = "Code for English Speaking", ylab = "Frequency",
        legend.text = rownames(tab.sexengl),
        beside = FALSE) # Stacked bars (default)
```

