---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

This page will contain all the assignments you submit for the class.




# Assignment 1

**Collaborators: Lorem Ipsum. **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.


```{r}
library(datasets)
```

**Answer**: I loaded the library! 

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

```{r}
dat<-USArrests
```

**Answer**: It is nice to rename data because we want to be able to replicate our analysis without messing up the original data.

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=FALSE}
dat$state <- tolower(rownames(USArrests))
```


List the variables contained in the dataset `USArrests`.

```{r}
names(dat)
```
**Answer**: The four variables contained in this dataset are Murder, Assault, UrbanPop and Rape.

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

**Answer**: Murder is a categorical variable, according to the DVB chapter.

What R Type of variable is it?

**Answer**: It is a character R Type variable. 


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

**Answer**: In this dataset we have the number of arrests per 100,000 residents for each of the following categories: assault, murder and rape in each of the 50 US states in 1973.

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

*Answer*: I chose to make a barplot instead of a histogram because it made more sense to me, considering that the data is categorical. I thought it would be clearer if each bar represented a state, which is why I decided to label the x-axis with the names of the states and the y-axis the number of arrests for murder per 100,000. In the code las=2 simply means that I wanted that state labels to be perpendicular to the x-axis.
```{r}
state.names = row.names(USArrests)
barplot(USArrests$Murder, names.arg = state.names, las = 2, ylab = "Arrest Rate for murder per 100,000", main = "Murder Rate in the United States in 1973")
```


### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
summary(dat$Murder)
```

**Answer**: Min: 0.800, 1st Q: 4.075, Median 7.250, Mean 7.788, 3rd Q 11.250, Max 17.400 ----- The mean is 7.788 and the median is 7.250. On average there were 7.788 (mean) arrests for murder per 100,000 residents in each state in 1973. The median tells us where the middle of the data set is. Since the mean is higher than the median we could say the distribution or arrests for murder per 100,000 in 1973 is positively skewed. A quartile represents 1/4 of the data, for examples the 1st Q is 4.075 and this tells us that 1/4 of the data lies below 4.075 arrests per 100,000. I think R only gives us the 1st and 3rd Quartiles because the median is equal to the 2nd Quartile. 

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r}
state.names = row.names(USArrests)
barplot(USArrests$Assault, names.arg = state.names, las = 2, ylab = "Arrest Rate for assault per 100,000", main = "Assault Rate in the United States in 1973")
```

```{r}
state.names = row.names(USArrests)
barplot(USArrests$Rape, names.arg = state.names, las = 2, ylab = "Arrest Rate for rape per 100,000", main = "Rape Rate in the United States in 1973")
```



```{r, echo = TRUE, fig.width = 5, fig.height = 8}

```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

**Answer**: Par can be used to set or query graphical parameters. Parameters can be set by specifying them as arguments to par in tag = value form, or by passing them as a list of tagged values.

What can you learn from plotting the histograms together?

**Answer**: By plotting histograms together (or in this case bar charts) we can see patterns more clearly. 
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval = FALSE, fig.width = 7.5, fig.height = 4}
library('maps') 
library('ggplot2') 

ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

Answer:

$$\\[2in]$$




# Assignment 2 

(Coming soon)
