---
title: "List of Codes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
summary(dat)
```


```{r}
names(dat)
```

# Dimension
```{r}
dim(dat)
```

1 variable bar plot.
```{r}
counts <- table(dat$age2)
barplot(counts, main = "Fig 1. Age of Participants", xlab="Code for Age")
```

Two variable stacked chart. 
```{r}
tab.agesex <- table(dat$irsex, dat$age2)
barplot(tab.agesex,
        main = "Fig 3. Relationship between Sex and Age",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.agesex),
        beside = FALSE) # Stacked bars (default)
```


Removing "99" or missing data 

dat$sexatract   \\\\ have 1 99 (missing data)
dat$sexatract.nas <-ifelse(dat$sexatract==99, NA, dat$seexatrct) 
table(dat$sexatract.nas)
table(dat$sexatract.nas)
na.omit(your.data.frame) 

Read(practicefatal-police-shootings-data.csv.htlm)


assignemnt 2 data 

dat$sexatract
dat$sexatract.nas <-ifelse(dat$sexatract==99, NA, dat$seexatract)

dat$sexatract.nas <- ifelse(dat$sexatract == 99, NA, dat$sexatract)

table(dat$sexatract.nas)

dat$sexatract.nas

sesxnona<-na.omit(dat$sexatract.nas)

table(dat$sexnona)

